---
title: "605_HW11.Rmd"
author: "Kumudini Bhave"
date: "April 18, 2017"
output:
  pdf_document: default
  html_document:
    fontsize: 35pt
    highlight: pygments
    theme: cerulean
---


# **LINEAR REGRESSION IN R **




### Problem Set 1 : 

### Using R's lm function, perform regression analysis and measure the significance of the independent variables for the following two data sets. In the first case, you are evaluating the statement that we hear that Maximum Heart Rate of a person is related to their age by the following equation:

$$MaxHR\quad =\quad 220\quad -\quad Age$$

The sample we have

Age | MaxHR
--- | ------
18 | 202 
23 | 186
25 | 187
35 | 180
65 | 156 
54 | 169
34 | 174
56 | 172
72 | 153
19 | 199
23 | 193
42 | 174
18 | 198
39 | 183
37 | 178


### Perform a linear regression analysis fitting the Max Heart Rate to Age using the lm function in R. What is the resulting equation? Is the effect of Age on Max HR significant? What is the significance level? Please also plot the fitted relationship between Max HR and Age.

### Solution :



```{r eval=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

# library for displaying data in tabular format
library(DT)

# library for plotting 
library(ggplot2)

age <- c(18, 23, 25, 35, 65, 54, 34, 56, 72, 19, 23, 42, 18, 39, 37)

maxhr <- c(202, 186, 187, 180, 156, 169, 174, 172, 153, 199, 193, 174, 198, 183, 178)

agehr <- data.frame(cbind(age, maxhr))

head(agehr)

# Taking the basic model of Max heart rate to Age
base.model <-  lm(maxhr ~ age, data = agehr)

summary(base.model)


```

From the summary we get the fitted line equation

$$\hat { y }\quad =\quad { \beta  }_{ 0 }\quad +\quad { \beta  }_{ 1 }x$$
as
$$\hat { maxhr } =\quad 210.04846\quad +\quad (-0.79773)\quad \times \quad age$$


**The Hypothesis Test**

${ H  }_{ 0 }$ : Age does not affect the max heart rate i.e. ${ \beta  }_{ 1 }$ = 0


${ H  }_{ A }$ : Age does affect the max heart rate i.e. ${ \beta  }_{ 1 }$ $\neq$ 0





From the summary we find the p value to be 3.85e-08 which is very much significant as it is much lower than 0.05
Hence age does seem to have a strong relationship with max heart rate for any individual.

Also we see that the coefficient for age is ${ \beta  }_{ 1 }$ $\neq$ 0 as ${ \beta  }_{ 1 }$ = -0.79773. The negative symbolizes that as age increases we find a decrease in the max heart rate for any individual.

Hence we reject the null hypothesis ${ H  }_{ 0 }$.

The fitted relationship between Maximum heart rate and Age can be plotted as below


```{r eval=TRUE,  tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

plotagehr <- ggplot(data=agehr, aes(y=maxhr, x=age)) + geom_point( colour="blue") + scale_x_continuous(name = "Age in Years") + scale_y_continuous(name="Max Heart Rate") + geom_abline(intercept = base.model$coefficients[1], slope = base.model$coefficients[2] , color = "red",  size = 1) + ggtitle("Max heart Rate Vs Age")


plotagehr

```






**********
### Problem Set 2 : 

### Using the Auto data set from Assignment 5 (also attached here) perform a Linear Regression analysis using mpg as the dependent variable and the other 4 (displacement, horsepower, weight, acceleration) as independent variables. What is the final linear regression fit equation? Which of the 4 independent variables have a significant impact on mpg? What are their corresponding significance levels? What are the standard errors on each of the coefficients? Please perform this experiment in two ways. First take any random 40 data points from the entire auto data sample and perform the linear regression fit and measure the 95% confidence intervals. Then, take the entire data set (all 392 points) and perform linear regression and measure the 95% confidence intervals. Please report the resulting fit equation, their significance values and confidence intervals for each of the two runs.   


### Solution :


 Extracting Raw Data From GitHub Data File And Reading In CSV Format

```{r eval=TRUE,  tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

# Loading RCurl package to help scrape data from web (stored on GitHub).
library(RCurl)

# Loading plyr package to help map abbreviated values to explained.
library(plyr)

data.giturl <- "https://raw.githubusercontent.com/DataDriven-MSDA/DATA605/master/auto-mpg.data"
autompg.data <-  read.table(data.giturl)
head(autompg.data)



# Attempt extract the required data columns of displacement, horsepower, weight, acceleration, mpg from the auto-mpg dataset.
autompg.datastudy  <- subset(autompg.data,select=c(V3,V4,V5,V6,V1), V4 != "?")
autompg.datastudy <- na.omit(autompg.datastudy)


# Verifying the number of attributes
length(autompg.datastudy)

# Verifying the number of observations selected
nrow(autompg.datastudy)

# Renaming the attributes
colnames(autompg.datastudy) <- c("disp", "hp", "wt", "acc","mpg")


# Viewing the data
head(autompg.datastudy)
dim(autompg.datastudy)
str(autompg.datastudy)

#Converting the horsepower as numeric as required , since it is quantitative variable
autompg.datastudy$hp <- as.numeric(autompg.datastudy$hp )
str(autompg.datastudy)


```


********

To understand the data and the relation between the various predictor variables (displacement, horsepower, weight , acceleration ) on  mpg, we plot the following scatterplots for mpg vs each of the predictor variables 

```{r eval=TRUE,  tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)


par(mfrow=c(1,2))


ggplot(data=autompg.datastudy, aes(y=mpg, x=disp)) + geom_point( colour="blue") + scale_x_continuous(name = "Displacement") + scale_y_continuous(name="MPG") + ggtitle("MPG Vs Displacement Scatterplot")



ggplot(data=autompg.datastudy, aes(y=mpg, x=hp)) + geom_point( colour="green") + scale_x_continuous(name = "Horsepower") + scale_y_continuous(name="MPG") + ggtitle("MPG Vs Horsepower Scatterplot")


par(mfrow=c(1,2))

ggplot(data=autompg.datastudy, aes(y=mpg, x=wt)) + geom_point( colour="purple") + scale_x_continuous(name = "Weight") + scale_y_continuous(name="MPG") + ggtitle("MPG Vs Weight Scatterplot")



ggplot(data=autompg.datastudy, aes(y=mpg, x=acc)) + geom_point( colour="red") + scale_x_continuous(name = "Acceleration") + scale_y_continuous(name="MPG") + ggtitle("MPG Vs Acceleration Scatterplot")


```

********
#### Auto Data Sample Study

Lets take the sample data from the whole population of auto data.

```{r eval=TRUE,  tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

# Function to randomly select samples from a given population dataset
# @param1 dataframe 
# @param2 number of samples 
rsamples <- function(df, num)
{
     set.seed(10)
     return (df[sample(nrow(df),num),])
     
}

# Taking 40 random samples form autodata population
autodatasamples <- rsamples(autompg.datastudy, 40)

head(autodatasamples)

dim(autodatasamples)
str(autodatasamples)


# Taking the model of mpg to all the predictors i.e. displacement, horsepower, weight, acceleration
auto.samplemodel <-  lm(mpg ~ ., data = autodatasamples)

samplesummary <- summary(auto.samplemodel)
samplesummary

```
From the summary we get the fitted line equation

$$\hat { mpg } \quad =\quad { \beta  }_{ 0 }\quad +\quad { \beta  }_{ 1 }\times disp\quad +\quad { \beta  }_{ 2 }\times hp\quad +\quad { \beta  }_{ 3 }\times wt\quad +\quad { \beta  }_{ 4 }\times acc$$

as
$$\hat { mpg } \quad =\quad 41.354484 \quad +\quad (-0.018761)\times disp\quad +\quad (0.024381)\times hp\quad +\quad (-0.005674)\times wt\quad +\quad (0.129268)\times acc$$

The  significance of the impact of the 4 independant variables on mpg can be gauged by the corresponding p values derived from the sample data, these are

Independant Variable |  p-value | Significance of Impact
-------------------- |--------- | ----------------------
displacement | 0.4648 | not significant
horsepower | 0.3598 | not significant
weight | 0.0361 | quite significant
accelaration | 0.6878 | not significant




The p-values for each of the independant variables   for sample data   are


```{r eval=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

samplesummary$coefficients[,4]


```

Form the p-values we derive that only weight seems to impact the mpg significantly. The displacement, horsepower, and acceleration do not have any significant impact on the mpg.



The standard error   for sample data  for the independant variables are

```{r eval=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

samplesummary$coefficients[,2]


```


The confidence interval   for sample data  can be calculated as

```{r eval=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

confint(auto.samplemodel, level=0.95)


```




********

#### Auto Data Population Study


Now considering the entire populaiton data for the autodata

```{r eval=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

dim(autompg.datastudy)
str(autompg.datastudy)

```


```{r eval=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)


#Taking the full model for the population 

auto.popmodel <-  lm(mpg ~ ., data = autompg.datastudy)

popsummary <- summary(auto.popmodel)
popsummary

```
From the summary we get the fitted line equation as

$$\hat { mpg } \quad =\quad 40.4449493 \quad +\quad ( -0.0102652)\times disp\quad +\quad 0.0064381\times hp\quad +\quad (-0.0061061)\times wt\quad +\quad 0.1828906\times acc$$




The  significance of the impact of the 4 independant variables on mpg can be gauged by the corresponding p values derived from the population data, these are

Independant Variable |  p-value | Significance of Impact
-------------------- |--------- | ----------------------
displacement | 0.1176 | not significant
horsepower | 0.4536 | not significant
weight | 4.58e-15 | quite significant
accelaration | 0.0631 | not significant


The p-values  for population data  for each of the independant variables are


```{r eval=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

popsummary$coefficients[,4]


```

As we observed with sample data, the population data for autodata also suggests that it is the weight that impacts the mpg significantly rather than the displacement, horsepower , acceleration.

The standard error for the independant variables for population data are

```{r eval=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

popsummary$coefficients[,2]


```


The confidence interval  for population data  can be calculated as


```{r eval=TRUE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)

confint(auto.popmodel, level=0.95)


```


We also observe that the standard error has significantly reduced from sample to population data since the 'n' of the data under study increases. This also results in a decrease in width of the confidence interval.


*******
