---
title: "605_HW8.Rmd"
author: "Kumudini Bhave"
date: "April 1, 2017"
output:
  pdf_document: default
  html_document:
    fontsize: 35pt
    highlight: pygments
    theme: cerulean
---

********


# **Bayes Theorem & Contingency Tables    **

********


### Problem Set 1 : 

### Your colleague either commutes by train or by the bus. 20 days of the month, she takes the train and the remaining 10 days she takes the bus. If she takes the train, she reaches work on time with a probability of 0:9. If she takes the bus, she frequently gets stuck in traffic and reaches work on time with a probability of 0:5. Given that she was on time today, what is the probability that she took the bus to work today?Your colleague either commutes by train or by the bus. 20 days of the month, she takes the train and the remaining 10 days she takes the bus. If she takes the train, she reaches work on time with a probability of 0:9. If she takes the bus, she frequently gets stuck in traffic and reaches work on time with a probability of 0:5. Given that she was on time today, what is the probability that she took the bus to work today?

### Solution :

Probability of taking bus = 10 / 30 = 1/3

Probability of taking train = 20/ 30 = 2/3

Probability of timely arrival with bus = 0.5

Probability of timely arrival with train = 0.9


Total days on time when taken train = 0.9 * 20 = 18 days
Total days on time when taken bus = 0.5 * 10 = 5 days


Total days on time = 18 + 5 = 23

Probability of the event of being on time , taken a bus is = 

(no of days on time when taken a bus )/total days on time 

5 / 23
=0.21739

Verifying with Bayes theorem

E : Event on time
X : Vehicle taken bus / train
X =x (taken bus)
X != x (taken train)

P(X = x|e) =   

          P(e|X = x) * P(X = x)  
          ______________________
                         P(e)


which is also

      P(e|X = x) * P(X = x) 
     _________________________ 
      P(e|X = x) * P(X = x) + P(e|X != x) * P(X != x)
 

     0.5 * (1/3)
     _________________________
     (0.5 * (1/3) + 0.9 * (2/3))

         0.16666 
     _________________________
        (0.16666 + 0.6)


= 0.16666/ 0.76666 =0.2174


**********


### Problem Set 2 : In the Grade Network that we looked at in the notes, what happens to the probability of Difficulty of Course when you present the evidence that the received recommendation letter was good? In addition, now present the evidence that both SAT scores were good and the letter of recommendation was good, What is the probability of the Difficulty of Course now? You should use the gRain package in R to build your network and perform these calculations. You may need to install RBGL package from BioConductor in R to get gRain working. See http://www.bioconductor.org/packages/release/bioc/html/RBGL.html for instructions on RBGL. Please submit your assignment as an R markdown document.


### Solution :

Enlisting the SAT grade parameters

We have level of **Difficulty**

Difficulty | Level | Pr
---------- | -------------
Yes | Y | 0.7
No | N | 0.3


We have level of **Intelligence**

Intelligence | Level
------------- | -------
High | H | 0.2
Low | L | 0.8


We have level of **Grade**

Difficulty / Intelligence | Level High | Level Low
------------------------- | ---------- | ----------
D=  No , I = Low  | 0.4 | 0.6
D = No, I = High | 0.99 | 0.01
D = Yes, I = Low | 0.2 | 0.8
D = Yes, I = High | 0.9 | 0.1


We have level of **SAT Score**

Intelligence | Level High | Level Low
------------ | ---------- | -----------
I = High | 0.8 | 0.2
I = Low | 0.1 | 0.9


We have level of **Letter Of Recommendation**

Grade | Level High | Level Low
----- | ---------- | ---------
G = High | 0.95 | 0.05 
G = Low | 0.1 | 0.9



```{r warning=FALSE, comment=FALSE, message=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=150)}
knitr::opts_chunk$set(message = FALSE, echo=TRUE)


# Library for graphical network of data
library(gRain)

```

```{r eval=TRUE}

#  Specifying conditional probability tables


# specifying levels
yn <- c("Yes","No")
hl <- c("High","Low")

d <- cptable(~Difficulty, values = c(0.7,0.3), levels = yn)
i <- cptable(~Intelligence, values = c(0.2,0.8), levels = hl)
#g.di <- cptable(~Grade|Difficulty : Intelligence, values = c(0.4, 0.6, 0.99, 0.01, 0.2, 0.8, 0.9, 0.1), levels = hl)
g.di <- cptable(~Grade|Difficulty : Intelligence, values = c( 0.9, 0.1, 0.99, 0.01, 0.2, 0.8, 0.4, 0.6), levels = hl)


s.i <- cptable(~SAT|Intelligence, values = c(0.8, 0.2, 0.1, 0.9), levels = hl)
l.g <- cptable(~Letter|Grade, values = c(0.95, 0.05, 0.1, 0.9), levels=hl)


# Compiliing list of conditional probability tables and forming network

 plist <- compileCPT(list(d, i, g.di, s.i, l.g))
 plist
gradenw <- grain(plist)
summary(gradenw)

# Plot model
 iplot(gradenw)
 
# Query model g
 querygrain(gradenw, nodes=c("Difficulty","Intelligence" , "Grade", "SAT", "Letter"))
 
```


```{r eval=TRUE}

plist$Difficulty

plist$Intelligence

plist$Grade

plist$SAT

plist$Letter



```



```{r eval=TRUE}



# Now we can add some data and repeat the query again.
# We try find that given Letter recommendation as High what is the probaility of Difficult of Course

gradenw_find1 <- setFinding(gradenw, nodes="Letter", states="High")
summary(gradenw_find1)


## Query model given parameter values (=prior) + some data
querygrain(gradenw_find1, nodes=c("Difficulty"))


```



```{r eval=TRUE}


# Now we can add some data and repeat the query again.
# We try find that given SAT score High, Letter recommendation as High,  what is the probaility of Difficult of Course

# First finding the SAT High stats
gradenw_find2 <- setFinding(gradenw, nodes="SAT", states="High")
summary(gradenw_find2)

# Finding the Letter High stats from the earlier findings
gradenw_find2 <- setFinding(gradenw_find2, nodes="Letter", states="High")
summary(gradenw_find2)

# Querying for probability for Difficulty , given SAT and Letter data from the latest findings
querygrain(gradenw_find2, nodes=c("Difficulty"))




```